---
editor_options: 
  chunk_output_type: console
---

# Model description

```{r}
library(data.table)
library(ggplot2)
library(patchwork)
library(colorspace)
```

![**Model implementation of discrete movement steps in continuous space, with movement steps selected based on inherited preferences for environmental cues.** In our model, **(A)** individuals search for food items (**green circles**), which may be immediately available (**filled green circles**; _F_), or may be available only in the future (**open green circles**). Individuals can sense only available items, and not unavailable ones. However, given our landscape structure, food items are clustered, making available items a good indicator of where resource clusters are (see next figure). Individuals can also sense other foraging individuals, and can sense whether they have successfully found, and are handling, a food item (handlers; **blue circles**), or whether they are unsuccessful foragers still searching for food (non-handlers; **filled grey circles**; _N_). To decide where to move, individuals sample their environment for these three cues (_F, H, N_) at 5 locations around themselves (**large open grey circles**), and have a sensory range of $d_S$. When the sensory range is relatively large (default = 1.0 units), there is some small overlap in samples. Individuals assign each potential direction a _suitability_, $S = s_FF + s_HH + s_NN + \epsilon$, where the coefficients $s_F, s_H, s_N$ are inherited preferences for environmental cues, and $\epsilon$ is a small error term that helps break ties between locations. In our implementation, the sensory distance ($d_S$) and the movement distance ($d_M$) are the same, 1.0 units. **(B)** Our infectious pathogen is transmitted between infected (**orange circles**) and susceptible (**filled grey circles**) individuals, with a probability $p$ = 0.05, when they are within a distance $d_\beta$ of each other. In our implementation, $d_\beta$ is the same as $d_S, d_M$ = 1.0 units.](figures/fig_schematic.png){ width=70% }

```{r}
repl = seq(4)
land = lapply(
  repl, function(x) {
    l = pathomove::get_test_landscape(
      nItems = 1800,
      landsize = 60,
      nClusters = 60,
      clusterSpread = 1,
      regen_time = 50
    )
    l$repl = x
    l
  }
)

land = rbindlist(land)
```

```{r}
# where agents are and where they sense
agent_sense = data.table(
  x = rep(5, 6),
  y = rep(5, 6),
  angle = c(0, seq(0, (2 * pi) - (2 * pi / 5), by = 2 * pi / 5))
)

agent_sense[, c("x", "y") := list(
  x + cos(angle),
  y + sin(angle) 
)]

# not first location
agent_sense[1, c("x", "y") := 5]

# assign type
agent_sense[, type := c("current", rep("potential", 5))]
```


```{r}
p_in =
  ggplot(land[repl == 3])+
  geom_point(
    aes(
      x, y,
    ),
    col = "forestgreen",
    size = 2
  )+
  ggforce::geom_circle(
    data = agent_sense,
    aes(
      x0 = x, y0 = y,
      r = 1,
      col = type
    ),
    alpha = 0.5,
    show.legend = FALSE
  )+
  geom_point(
    x = 5, y = 5,
    col = "black",
    shape = 17,
    size = 3
  )+
  scale_colour_discrete_diverging(
    palette = "Blue-Red 2",
    rev = F
  )+
  theme_test()+
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text = element_blank(),
    strip.background = element_blank(),
    plot.background = element_blank(),
    panel.background = element_rect(
      fill = "grey99"
    )
  )+
  coord_equal(
    xlim = c(0, 10),
    ylim = c(0, 10)
  )

p_out = ggplot(land[repl == 3])+
  geom_point(
    aes(
      x, y,
    ),
    col = "forestgreen",
    size = 0.2
  )+
  ggforce::geom_circle(
    data = agent_sense,
    aes(
      x0 = x, y0 = y,
      r = 1,
      col = type
    ),
    alpha = 0.5,
    show.legend = FALSE
  )+
  geom_point(
    x = 5, y = 5,
    col = "black",
    shape = 17,
    size = 3
  )+
  scale_colour_discrete_diverging(
    palette = "Blue-Red 2",
    rev = F
  )+
  annotation_custom(
    grob = ggplotGrob(
      p_in
    ),
    xmin = 30, xmax = 60,
    ymin = 30, ymax = 60
  )+
  theme_test()+
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    strip.background = element_blank()
  )+
  coord_equal()

ggsave(
  p_out,
  height = 120, width = 120,
  units = "mm",
  filename = "supplement/figures/fig_landscape.png"
)
```

![**An example of the resource landscape used in our simulations.** Our simulation's resource landscape consists of 60 randomly distributed clusters of food items ('resource patches'), with 1800 discrete food items divided among the clusters (30 items per cluster). The landscape is a square of 60 units per side, with wrapped boundaries (i.e., a torus). The food item density in our scenarios is 0.5 food items per unit area. Items are distributed around the centre of each cluster, within a standard deviation of 1.0 unit. Items, once consumed by foragers, are unavailable for a fixed number of timesteps (the regeneration time $R$, expressed in terms of the foragers' generation time), after which they regenerate in the same location. While regenerating (i.e., unavailable). While regenerating, items cannot be sensed by foragers. The sensory ranges of individuals ($d_S$) are shown for each potential step (**red circles**, including the current location: **blue circle**). Food item clustering means that available items, as well as foragers handling a food item (handlers) are good indicators of the location of a resource cluster.](figures/fig_landscape.png){ width=70% }

\newpage
