---
editor_options:
  chunk_output_type: console
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## Sporadic introduction of infectious pathogens

We implemented a variant of our main model, in which the infectious pathogen is introduced only sporadically after the first introduction event (at G = 3,000).
Specifically, we modelled probabilistic introduction of the pathogen in each generation following the initial introduction.
We call the per-generation probability of a novel pathogen introduction event the 'spillover rate'.
We ran 10 replicates each of this model variant and examined whether there was a similar evolutionary shift in social movement strategies as seen in our default implementation.
Since it is the main parameter of interest, we ran this model variant for three values of the spillover rate: 0.05, 0.1, and 0.25.
Instead of examining the joint effect of landscape productivity and cost of infection as well, we only examined the effect of infection cost, implementing three different variants with an infection cost $\delta E$ of 0.1, 0.25, and 0.5.
We kept all other model parameters similar to the default scenario of our main model, and importantly, considered only a landscape productivity $R$ of 2.
Cross-species novel pathogen introductions are likely to become more common with climate change, and so we chose these spillover rate values to represent different scenarios under altered global regimes of pathogen transfer.
Our model's default implementation may be seen as an extreme case of the models considered here, with a spillover rate of 1.0.

### Model implementation {-}

In our model code, the sporadic introduction is implemented by drawing the number of generations until the next pathogen introduction event from a geometric distribution whose probability parameter is given by the spillover rates described above.
Zero values are handled by converting them into ones.
At our lowest spillover rate, up to 100 generations could pass between pathogen introductions, while at our highest rates, there are rarely more than 10 generations between introductions.

```{r eval=FALSE,include=FALSE}
# load libraries
library(data.table)
library(ggplot2)
library(colorspace)
library(patchwork)
```


```{r eval=FALSE, include=FALSE}
files = list.files(
  "data/results/morph_data", 
  pattern = "csv",
  full.names = TRUE
)

data_all = lapply(files, fread)
df_strat = rbindlist(data_all)

# get sim type
df_strat = df_strat[scenario == 3,]

# popsize
popsize = 500
replicates = 10
df_strat[, prop := N / popsize]
g_bin = 100
df_strat[, gen_bin := (floor(gen / g_bin) * g_bin)]

# get time since pathogen
sgen = 3000
genmax = 5000
df_strat[, gen_abs := gen_bin - sgen]

df_strat = df_strat[gen %% 100 == 0,]

df_strat = df_strat[, list(
  prop = sum(prop) / (replicates)
), by = c(
  "gen", "social_strat", 
  "cost", "regen", "dispersal", "spillover_rate"
)]

# regeneration rate
gen_time = 100
df_strat[, regen_r := gen_time / regen]
```

```{r eval=FALSE, include=FALSE}
fig_evo_strategy_sporadic = 
  ggplot(df_strat)+
  geom_col(
    aes(
      gen, prop,
      fill = social_strat
    ),
    width = 100,
    position = "stack"
  )+
  geom_vline(
    xintercept = sgen,
    lty = 2,
    size = 0.3,
    col = "red"
  )+
  scale_fill_discrete_sequential(
    palette = "Viridis",
    l2 = 80,
    rev = F,
    name = NULL,
    limits = c("agent avoiding", "agent tracking", "handler tracking"),
    labels = stringr::str_to_sentence,
    na.value = "lightgrey"
  )+
  scale_x_continuous(
    breaks = c(1000, sgen, genmax),
    name = "Generations",
    labels = scales::comma,
    sec.axis = dup_axis(
      breaks = sgen,
      labels = "Initial pathogen introduction",
      name = "Increasing probability of pathogen introduction \u279c"
    )
  )+
  scale_y_continuous(
    labels = scales::percent,
    breaks = NULL,
    name = glue::glue("Increasing disease cost \u279c"),
    sec.axis = dup_axis(
      breaks = c(0, 0.5, 1),
      labels = scales::percent,
      name = "% Individuals"
    )
  )+
  facet_grid(
    cost ~ spillover_rate,
    as.table = F, 
    switch = c("y"),
    labeller = labeller(
      spillover_rate = function(x) {
        scales::percent(as.numeric(x), prefix = "p(Intro) = ")
      },
      cost = function(x) {
        if(x >= 0.1) {
          sprintf("δE = %s", x)
        } else {
          scales::percent(as.numeric(x), prefix = "δE = ")
        }
      }
    )
  )+
  coord_cartesian(
    expand = F
  )+
  theme_test(
    base_size = 8, 
    base_family = "Arial"
  )+
  theme(
    legend.position = "top", 
    legend.key.height = unit(1, "mm"),
    strip.placement = "outside",
    strip.text = ggtext::element_markdown(),
    axis.text.x = element_text(hjust = 0.5, size = 6),
    axis.text.x.top = element_text(
      colour = "red"
    )
  )

# plot generations between introductions
p = c(0.05, 0.1, 0.25)
gen_intervals = lapply(p, rgeom, n = 1000)
data_gen = data.table(
  p = p, gen_intervals
)
# unnest
data_gen = data_gen[, list(
  gen_intervals = unlist(gen_intervals)
), by = c("p")]
# add one
data_gen[, gen_intervals := gen_intervals + 1]

# figure of generations between introductions
fig_gen_interval = 
  ggplot(data_gen)+
  geom_histogram(
    aes(gen_intervals, y = ..count.. / 1000),
    binwidth = 2,
    fill = "steelblue"
  )+
  facet_grid(
    cols = vars(p),
    as.table = F, 
    switch = c("y"),
    labeller = labeller(
      p = function(x) {
        scales::percent(as.numeric(x), prefix = "p(Intro) = ")
      }
    )
  )+
  coord_cartesian(
    xlim = c(1, 50),
    expand = F
  )+
  scale_y_continuous(
    labels = scales::percent,
    name = "Probability"
  )+
  scale_x_continuous(
    name = "Gens. between pathogen introductions"
  )+
  theme_test(
    base_size = 8, 
    base_family = "Arial"
  )+
  theme(
    legend.position = "top", 
    legend.key.height = unit(1, "mm"),
    strip.placement = "outside",
    strip.text = ggtext::element_markdown(),
    axis.text.x = element_text(hjust = 0.5, size = 6),
    axis.text.x.top = element_text(
      colour = "red"
    )
  )

# wrap figures
fig_evo_strategy_sporadic =
wrap_plots(
  fig_gen_interval,
  fig_evo_strategy_sporadic,
  ncol = 1,
  design = "A\nB\nB\nB"
) &
  plot_annotation(
    tag_levels = c("A")
  ) &
  theme(
    plot.tag = element_text(
      face = "bold"
    )
  )

# save figure
ggsave(
  plot = fig_evo_strategy_sporadic,
  filename = "supplement/figures/fig_evo_strategy_sporadic.png",
  height = 160, width = 120, units = "mm"
)
```

### Evolutionary outcomes of the percentage cost implementation

The social movement strategies evolved prior to pathogen introduction are identical to those seen in our default implementation, as expected.
However, following pathogen introduction, we found that there was little to change in the population-level mixture of movement strategies in this model variant (see figure).
This is regardless of the probability of a novel pathogen introduction (our so-called 'spillover rate'), and the cost of infection by a pathogen.
Across the simulation, the commonest social movement strategy remains 'agent tracking', i.e., preferring locations with multiple individuals regardless of their foraging status.

Since there is little to no change in social movement strategies, we did not expect nor find changes in ecological outcomes.

![**No evolutionary change in social movement strategies when novel pathogen introduction events are relatively uncommon.** **(A)** In our alternative implementation of the model, the pathogen is only introduced sporadically after the initial introduction (G = 3,000; red line in panel B). **(B)** When the introductions are relatively rare and sporadic, there is no shift in the mixture of movement strategies after pathogen introduction. The agent tracking strategy remains common across parameter combinations. ](figures/fig_evo_strategy_sporadic.png){ width=70% }
