---
editor_options: 
  chunk_output_type: console
---

# Script to send simulation jobs to the HPC cluster


```{r}
# load libraries
library(pathomove)
library(stringr)
library(glue)

password <- readLines("data/password")
```

## Scenario 1: Repeated introduction

### Default case with repeated introduction

```{r}
# param file name
date <- Sys.time() |> str_replace_all(" |:", "_")

scenario_tag <- "default"
# prepare parameters for default case
param_file_default <- glue("data/parameters/parameters_{date}_{scenario_tag}.csv")

pathomove::make_parameter_file(
  scenario = 1,
  popsize = 500,
  nItems = 1800,
  landsize = 60,
  nClusters = 60,
  clusterSpread = 1,
  tmax = 100,
  genmax = 5000,
  g_patho_init = 3000,
  range_food = 1,
  range_agents = 1,
  range_move = 1,
  handling_time = 5,
  regen_time = c(20, 50, 100),
  pTransmit = 0.05,
  initialInfections = 20,
  costInfect = c(0.1, 0.25, 0.5),
  multithreaded = FALSE,
  replicates = 10,
  dispersal = 2.0, # for local-ish dispersal
  vertical = FALSE,
  infect_percent = FALSE,
  mProb = 0.01,
  mSize = 0.01,
  spillover_rate = 0.0,
  which_file = param_file_default
)

# try sending in a job
pathomove::use_cluster(
  ssh_con = "p284074@peregrine.hpc.rug.nl",
  password = password,
  script = "scripts/do_sim_pathomove.R",
  tag = scenario_tag,
  folder = "patho-move-evol",
  template_job = "bash/main_job_maker_default.sh",
  parameter_file = param_file_default
)
```

### Repeated introduction with global natal dispersal

```{r}
# param file name
date <- Sys.time() |> str_replace_all(" |:", "_")

scenario_tag <- "global"
# prepare parameters for global dispersal case
param_file_global <- glue("data/parameters/parameters_{date}_{scenario_tag}.csv")

pathomove::make_parameter_file(
  scenario = 1,
  popsize = 500,
  nItems = 1800,
  landsize = 60,
  nClusters = 60,
  clusterSpread = 1,
  tmax = 100,
  genmax = 5000,
  g_patho_init = 3000,
  range_food = 1,
  range_agents = 1,
  range_move = 1,
  handling_time = 5,
  regen_time = c(20, 50, 100),
  pTransmit = 0.05,
  initialInfections = 20,
  costInfect = c(0.25),
  multithreaded = FALSE,
  replicates = 10,
  dispersal = 10.0, # for global-ish dispersal
  # vertical = FALSE,
  infect_percent = FALSE,
  mProb = 0.01,
  mSize = 0.01,
  # spillover_rate = 0.01,
  which_file = param_file_global
)

# try sending in a job
pathomove::use_cluster(
  ssh_con = "p284074@peregrine.hpc.rug.nl",
  password = password,
  script = "scripts/do_sim_pathomove.R",
  tag = scenario_tag,
  folder = "patho-move-evol",
  template_job = "bash/main_job_maker_global.sh",
  parameter_file = param_file_global
)
```

### Repeated introduction with percentage costs

```{r}
# param file name
date <- Sys.time() |> str_replace_all(" |:", "_")

scenario_tag <- "percent"
# prepare parameters for percentage costs case
param_file_percent <- glue("data/parameters/parameters_{date}_{scenario_tag}.csv")

pathomove::make_parameter_file(
  scenario = 1,
  popsize = 500,
  nItems = 1800,
  landsize = 60,
  nClusters = 60,
  clusterSpread = 1,
  tmax = 100,
  genmax = 5000,
  g_patho_init = 3000,
  range_food = 1,
  range_agents = 1,
  range_move = 1,
  handling_time = 5,
  regen_time = c(20, 50, 100),
  pTransmit = 0.05,
  initialInfections = 20,
  costInfect = c(0.025, 0.05, 0.075),
  multithreaded = FALSE,
  replicates = 10,
  dispersal = 2.0, # for local-ish dispersal
  # vertical = FALSE,
  infect_percent = TRUE,
  mProb = 0.01,
  mSize = 0.01,
  # spillover_rate = 0.01,
  which_file = param_file_percent
)

# try sending in a job
pathomove::use_cluster(
  ssh_con = "p284074@peregrine.hpc.rug.nl",
  password = password,
  script = "scripts/do_sim_pathomove.R",
  tag = "percent",
  folder = "patho-move-evol",
  template_job = "bash/main_job_maker_percent.sh",
  parameter_file = param_file_percent
)
```

### Repeated introduction with reproduction energy threshold

```{r}
# param file name
date <- Sys.time() |> str_replace_all(" |:", "_")

scenario_tag <- "threshold"
# prepare parameters for reproduction threshold case
param_file_threshold <- glue("data/parameters/parameters_{date}_{scenario_tag}.csv")

pathomove::make_parameter_file(
  scenario = 1,
  popsize = 500,
  nItems = 1800,
  landsize = 60,
  nClusters = 60,
  clusterSpread = 1,
  tmax = 100,
  genmax = 5000,
  g_patho_init = 3000,
  range_food = 1,
  range_agents = 1,
  range_move = 1,
  handling_time = 5,
  regen_time = c(50),
  pTransmit = 0.05,
  initialInfections = 20,
  costInfect = c(0.1, 0.25, 0.5),
  multithreaded = FALSE,
  replicates = 10,
  dispersal = 2.0, # for global-ish dispersal
  # vertical = FALSE,
  infect_percent = FALSE,
  reprod_threshold = TRUE,
  mProb = 0.01,
  mSize = 0.01,
  # spillover_rate = 0.01,
  which_file = param_file_threshold
)

# try sending in a job
pathomove::use_cluster(
  ssh_con = "p284074@peregrine.hpc.rug.nl",
  password = password,
  script = "scripts/do_sim_pathomove.R",
  folder = "patho-move-evol",
  tag = "threshold",
  template_job = "bash/main_job_maker_threshold.sh",
  parameter_file = param_file_threshold
)
```

### Repeated but sporadic introduction

```{r}
# param file name
date <- Sys.time() |> str_replace_all(" |:", "_")

scenario_tag <- "sporadic"
# prepare parameters for sporadic introduction case
param_file_sporadic <- glue("data/parameters/parameters_{date}_{scenario_tag}.csv")

pathomove::make_parameter_file(
  scenario = 3,
  popsize = 500,
  nItems = 1800,
  landsize = 60,
  nClusters = 60,
  clusterSpread = 1,
  tmax = 100,
  genmax = 5000,
  g_patho_init = 3000,
  range_food = 1,
  range_agents = 1,
  range_move = 1,
  handling_time = 5,
  regen_time = c(50),
  pTransmit = 0.05,
  initialInfections = 20,
  costInfect = c(0.1, 0.25, 0.5),
  multithreaded = FALSE,
  replicates = 10,
  dispersal = 2.0, # also local-ish dispersal
  vertical = FALSE,
  infect_percent = FALSE,
  mProb = 0.01,
  mSize = 0.01,
  spillover_rate = c(0.1, 0.25, 0.5),
  which_file = param_file_sporadic
)

# try sending in a job
pathomove::use_cluster(
  ssh_con = "p284074@peregrine.hpc.rug.nl",
  password = password,
  script = "scripts/do_sim_pathomove.R",
  folder = "patho-move-evol",
  template_job = "bash/main_job_maker_sporadic.sh",
  parameter_file = param_file_sporadic
)
```

### Repeated introduction with handling time alternatives

```{r}
# param file name
date <- Sys.time() |> str_replace_all(" |:", "_")

password <- readLines("data/password")

scenario_tag <- "handling"

# prepare parameters for percent case
param_file_handling <- glue("data/parameters/parameters_{date}_{scenario_tag}.csv")

pathomove::make_parameter_file(
  scenario = 1,
  popsize = 500,
  nItems = 1800,
  landsize = 60,
  nClusters = 60,
  clusterSpread = 1,
  tmax = 100,
  genmax = 5000,
  g_patho_init = 3000,
  range_food = 1,
  range_agents = 1,
  range_move = 1,
  handling_time = c(0, 1, 2, 10),
  regen_time = 50,
  pTransmit = 0.05,
  initialInfections = 20,
  costInfect = 0.25,
  multithreaded = FALSE,
  replicates = 10,
  dispersal = 2.0, # for local-ish dispersal
  # vertical = FALSE,
  infect_percent = FALSE,
  mProb = 0.01,
  mSize = 0.01,
  # spillover_rate = 0.01,
  which_file = param_file_handling
)

# try sending in a job
pathomove::use_cluster(
  ssh_con = "p284074@peregrine.hpc.rug.nl",
  password = password,
  script = "scripts/do_sim_pathomove.R",
  folder = "patho-move-evol",
  tag = scenario_tag,
  template_job = "bash/main_job_maker_handling.sh",
  parameter_file = param_file_handling
)
```

### Repeated introduction with alternative spatial structure

```{r}
# param file name
date <- Sys.time() |> str_replace_all(" |:", "_")

password <- readLines("data/password")

scenario_tag <- "spatial"

# prepare parameters for percent case
param_file_spatial <- glue("data/parameters/parameters_{date}_{scenario_tag}.csv")

pathomove::make_parameter_file(
  scenario = 1,
  popsize = 500,
  nItems = 1800,
  landsize = 60,
  nClusters = 10,
  clusterSpread = c(1, 10),
  tmax = 100,
  genmax = 5000,
  g_patho_init = 3000,
  range_food = 1,
  range_agents = 1,
  range_move = 1,
  handling_time = 5,
  regen_time = 50,
  pTransmit = 0.05,
  initialInfections = 20,
  costInfect = 0.25,
  multithreaded = FALSE,
  replicates = 10,
  dispersal = 2.0, # for local-ish dispersal
  # vertical = FALSE,
  infect_percent = FALSE,
  mProb = 0.01,
  mSize = 0.01,
  # spillover_rate = 0.01,
  which_file = param_file_spatial
)

# try sending in a job
pathomove::use_cluster(
  ssh_con = "p284074@peregrine.hpc.rug.nl",
  password = password,
  script = "scripts/do_sim_pathomove.R",
  folder = "patho-move-evol",
  tag = scenario_tag,
  template_job = "bash/main_job_maker_spatial.sh",
  parameter_file = param_file_spatial
)
```

## Scenario 2: Single introduction with vertical transmission

```{r}
# param file name
date <- Sys.time() |> str_replace_all(" |:", "_")

scenario_tag <- "vertical"
# prepare parameters for reproduction threshold case
param_file_vertical <- glue("data/parameters/parameters_{date}_{scenario_tag}.csv")

pathomove::make_parameter_file(
  scenario = 2,
  popsize = 500,
  nItems = 1800,
  landsize = 60,
  nClusters = 60,
  clusterSpread = 1,
  tmax = 100,
  genmax = 5000,
  g_patho_init = 3000,
  range_food = 1,
  range_agents = 1,
  range_move = 1,
  handling_time = 5,
  regen_time = c(50),
  pTransmit = 0.05,
  p_v_transmit = c(0.1, 0.2, 0.3),
  initialInfections = 20,
  costInfect = c(0.1, 0.25, 0.5),
  multithreaded = FALSE,
  replicates = 10,
  dispersal = 2.0, # also local-ish dispersal
  vertical = TRUE,
  infect_percent = FALSE,
  mProb = 0.01,
  mSize = 0.01,
  which_file = param_file_vertical
)

# try sending in a job
pathomove::use_cluster(
  ssh_con = "p284074@peregrine.hpc.rug.nl",
  password = password,
  script = "scripts/do_sim_pathomove.R",
  folder = "patho-move-evol",
  tag = "vertical",
  template_job = "bash/main_job_maker_vertical.sh",
  parameter_file = param_file_vertical
)
```
