---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Plot networks in final generations

```{r}
library(data.table)
library(glue)
library(tidygraph)
library(ggraph)
library(colorspace)
library(patchwork)
```

## Load networks from default scenario

Read parameter files to subset for default parameter combination, based on the "ofi" column.

```{r}
params = list.files(
  "data/parameters", pattern = "csv", full.names = T
)
params = lapply(params, fread) |>
  rbindlist(use.names = TRUE)
```

```{r}
params = params[scenario == 1 & ((!infect_percent) & 
                  (costInfect == 0.25) & 
                  (regen_time == 50) & 
                  (dispersal == 2))]

data_files = glue("data/results/networks/data_networks_{params$ofi}.Rds")
```

```{r}
data_ntwk = lapply(data_files, function(file) {
  nt = readRDS(file)
  nt
})
```

## Degree distribution

```{r}
degree_data = lapply(
  data_ntwk, function(le) {
    le_pre = le[["3000"]]
    le_post = le[["3500"]]
    # pre pathogen degree
    le_pre = mutate(
      le_pre, degree = tidygraph::centrality_degree(
        normalized = F
      ),
      type = "pre"
    ) |>
      as_tibble()
    
    # post pathogen degree
    le_post = mutate(
      le_post, degree = tidygraph::centrality_degree(
        normalized = F
      ),
      type = "post"
    ) |>
      as_tibble()
    
    rbindlist(
      list(
        le_pre,
        le_post
      )
    )
  }
) |>
  rbindlist()

# save degree data
fwrite(
  degree_data,
  "data/results/data_default_degree_distribution.csv"
)
```

```{r}
degree_data = fread("data/results/data_default_degree_distribution.csv")
degree_data$type = factor(degree_data$type, levels = c("pre", "post"))
```

```{r}
# popsize per sim
popsize = 500
# how many replicates
replicates = 10

fig_degree =
ggplot(degree_data)+
  geom_histogram(
    aes(
      degree,
      fill = type,
      y = ..count.. / (popsize * replicates)
    ),
    position = "identity",
    bins = 15,
    col = NA,
    show.legend = F
  )+
  geom_vline(
    xintercept = 0.1 * popsize,
    lty = 2,
    size = 0.2
  )+
  scale_fill_discrete_diverging(
    palette = "Blue-Red",
    rev = F,
    l1 = 50
  )+
  scale_x_continuous(
    trans = ggallin::pseudolog10_trans,
    breaks = c(0, 10, 50, 500),
    labels = function(x) {
      scales::percent(
        accuracy = 1,
        as.numeric(x) / popsize
      )
    },
    name = "% Pop. encountered"
  )+
  scale_y_continuous(
    # breaks = c(0, 0.1, 0.25),
    labels = scales::percent_format(accuracy = 1)
  )+
  theme_test(
    base_family = "Arial",
    base_size = 8
  )+
  theme(
    legend.position = "top",
    legend.key.height = unit(0.5, "mm"),
    legend.key.width = unit(2, "mm"),
    axis.text.y = element_text(
      angle = 90,
      hjust = 0.5,
      size = 6
    ),
    axis.text.x = element_text(
      size = 6
    ),
    plot.background = element_blank(),
    panel.background = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(
      face = "bold"
    )
  )+
  labs(
    y = "% Indiv.",
    title = "Inset 1"
  )+
  guides(
    fill = guide_legend()
  )
```

## Default scenario network plots

```{r}
# select nice network
ntwks_example = data_ntwk[[5]]

# select before and after disease
ntwk_pre = ntwks_example[["2500"]] %>% 
  activate(edges) %>% 
  filter(weight > quantile(weight, probs = 0.25)) %>% 
  activate(nodes)

# sanity check
ggraph(ntwk_pre, x = xn, y = yn)+
  geom_node_point(
    aes(
      fill = t_infec,
      size = assoc
    ),
    shape = 21,
    show.legend = F
  )+
  coord_equal(
    xlim = c(0, 60),
    ylim = c(0, 60)
  )

# select a nice network. Note this may be a different replicate
# networks are plotted for illustration only
ntwks_example = data_ntwk[[3]]
ntwk_post = ntwks_example[["3500"]] %>%
  activate(edges) %>%
  filter(weight > quantile(weight, probs = 0.25)) %>%
  activate(nodes)
```

```{r}
# make network figures
networkplots = lapply(
  list(ntwk_pre, ntwk_post), function(n) {
    ggraph(n, x = xn, y = yn)+
      geom_edge_fan(
        edge_width = 0.5,
        aes(
          edge_alpha = weight
        ),
        edge_color = "grey40",
        show.legend = F
      )+
      geom_node_point(
        aes(
          fill = t_infec,
          size = assoc
        ),
        shape = 21,
        show.legend = T
      )+
      scale_size_continuous(
        range = c(0.5, 3)
      )+
      scale_fill_continuous_sequential(
        palette = "Inferno",
        limit = c(1, 100),
        breaks = c(1, 10, 30, 100),
        # direction = -1,
        na.value = "lightblue",
        trans = "sqrt"
      )+
      coord_cartesian(
        expand = TRUE,
        xlim = c(0, 60),
        ylim = c(0, 60)
      )+
      theme_graph(
        background = "white",
        border = T,
        base_size = 8,
        plot_margin = margin(rep(0, 3))
      )+
      theme(
        # axis.ticks = element_blank(),
        # axis.text = element_blank(),
        # axis.title = element_blank(),
        legend.margin = margin(rep(0, 4)),
        legend.position = "top",
        legend.title = element_text(size = 6),
        legend.key.height = unit(1, units = "mm"),
        legend.key.width = unit(3, units = "mm"),
        plot.background = element_blank()
      )+
      labs(
        fill = "Time infected"
      )+
      guides(
        size = "none",
        edge_alpha = "none"
      )
  }
)

# select the inset graph -- this is hardcoded
ntwk_inset = ntwk_pre %>% 
  activate(edges) %>% 
  filter(weight > quantile(weight, probs = 0.9)) %>% 
  activate(nodes) %>% 
  filter(
    xn > 15, xn < 25,
    yn > 45, yn < 55
  )

# make network inset
ntwk_inset_plot = ggraph(ntwk_inset, x = xn, y = yn)+
  geom_edge_fan(
    edge_width = 0.1,
    edge_color = alpha("grey", 0.3),
    show.legend = F
  )+
  geom_node_point(
    # size = 3,
    fill = "lightblue",
    shape = 21,
    # colour = "grey50",# "grey50",
    show.legend = F
  )+
  coord_cartesian(
    expand = F,
    xlim = c(18, 20),
    ylim = c(50, 52)
  )+
  theme_graph(
    background = "white",
    border = T,
    base_size = 8,
    plot_margin = margin(rep(0, 3))
  )+
  theme(
    plot.background = element_rect(
      fill = "grey99",
      colour = "indianred",
      size = 1
    )
  )
  
# add degree distribution to panel A
networkplots[[1]] =
  networkplots[[1]] +
  annotate(
    geom = "rect",
    xmin = 15, xmax = 25,
    ymin = 45, ymax = 55,
    fill = "transparent",
    col = "indianred",
    size = 0.5
  )+
  annotation_custom(
    grob = ggplot2::ggplotGrob(fig_degree),
    xmin = 20,
    xmax = 60,
    ymin = 0,
    ymax = 40
  )+
  annotate(
    geom = "segment",
    x = 25, xend = 40,
    y = 50, yend = 50,
    col = "indianred"
  )+
  annotation_custom(
    grob = ggplot2::ggplotGrob(ntwk_inset_plot),
    xmin = 40,
    xmax = 60,
    ymin = 40,
    ymax = 60
  )
```

## Plot SIR model runs

```{r}
# spillover data and rounding
popsize = 500
replicates = 10
data_sir_models = fread("data/results/data_sir_models.csv")

# select scenarios
data_sir_models = data_sir_models[scenario == "novel pathogen"]

# data_sir_models$mean = data_sir_models$mean / 500
data_sir_models = data_sir_models[class == "NR" & gamma == 1 &
                                    beta == 5 & threshold %in% c(1, 10) ]

# select scenario
data_default = data_sir_models[regen_time == 50 & costInfect == 0.25]

# round time, with larger bins later in time
# ie, increments of 0, 0.25, 0.5, 1, 2, 5, 10 etc.
data_default[, tround := plyr::round_any(time, 0.25)]
data_default[, tround := fifelse(
  time > 5, plyr::round_any(time, 0.5),
  tround
)]
data_default[, type := factor(type, levels = c("pre", "post"))]

# save data default sir
fwrite(
  data_default,
  file = "data/results/data_default_sir.csv"
)
```

```{r}
# load saved data to save time
data_default = fread("data/results/data_default_sir.csv")
data_default[, type := factor(type, levels = c("pre", "post"))]
# data_default = split(data_default, by = "threshold")
```

```{r}
# plot the progress of disease in default network sir models
fig_sir =
  ggplot(data_default)+
    geom_hline(
      yintercept = 0.5,
      col = "grey",
      size = 0.2
    )+
    geom_bin2d(
      aes(
        time, mean / popsize,
        fill = type,
        alpha = ..density..
      ),
      show.legend = F
    )+
    stat_summary(
      aes(
        tround, mean / popsize, 
        colour = type,
        shape = type,
        group = interaction(type, threshold, beta)
      ),
      stroke = 0.8,
      size = 0.3,
      position = position_dodge(width = 0.1),
      show.legend = F
    )+
    facet_wrap(
      ~threshold,
      labeller = labeller(
        threshold = c(
          "1" = "1 encounter",
          "10" = "10 encounters"
        )
      )
    )+
    scale_colour_discrete_diverging(
      l2 = 50,# l2 = 40,
      palette = "Blue-Red 2",
      rev = FALSE,
      # palette = "Plasma",
      breaks = c("pre", "post"),
      labels = c("Pre-pathogen intro.", "Post-pathogen intro.")
    )+
    scale_fill_discrete_diverging(
      palette = "Blue-Red 2",
      rev = FALSE,
      # palette = "Plasma",
      breaks = c("pre", "post"),
      labels = c("Pre-pathogen intro.", "Post-pathogen intro.")
    )+
    scale_shape_manual(
      values = c(
        "pre" = 16,
        "post" = 17
      ),
      breaks = c("pre", "post"),
      labels = c("Pre-pathogen intro.", "Post-pathogen intro.")
    )+
    scale_y_continuous(
      labels = scales::percent,
      breaks = c(0, 0.5, 1)
      # breaks = c(0.01, 0.33, 0.67, 1)
    )+
    # scale_x_sqrt()+
    theme_test(
      base_size = 8
    )+
    theme(
      legend.position = "top",
      strip.background = element_blank(),
      strip.text = element_text(
        face = "bold"
      ),
      # strip.placement = "outside",
      axis.text.y = element_text(
        angle = 90,
        hjust = 0.5,
        size = 6
      )
    )+
    coord_cartesian(
      xlim = c(0, 6),
      ylim = c(0, 1)
    )+
    labs(
      x = "SIR model time",
      y = "% Agents infected",
      fill = NULL,
      colour = NULL,
      shape = NULL
    )
```

## Infections per generation

```{r}
# some parameters
sgen = 3000
genmax = 5000
popsize = 500

# load data
data_infections_gen = fread("data/results/data_infections_gen.csv")

# subset data
data_infections_gen = data_infections_gen[scenario == 1]
data_infections_gen = data_infections_gen[cost == 0.25 & regen == 50 
                      & dispersal == 2]
# data_infections_gen[, gen_bin := round(gen, digits = -2)]

# absolute generations
data_infections_gen[, gen_abs := gen - sgen]

# make smaller bins
# data_infections_gen[, gen_abs := fifelse(
#   abs(gen_abs) < 100, gen - sgen, gen_abs
# )]
```

```{r}
# make figures
fig_gen_infect =
ggplot(data_infections_gen[gen >= sgen & gen %% 10 == 0])+
  geom_vline(
    xintercept = c(0, 25),
    lty = 2, 
    size = 0.2,
    col = c("red", "grey40")
  )+
  geom_bin2d(
    aes(
      gen_abs, n_infected / popsize
    ),
    binwidth = c(1, 0.01),
    show.legend = F
  )+
  stat_summary(
    aes(
      gen_abs, n_infected / popsize
    ),
    col = "brown",
    geom = "line",
    show.legend = FALSE
  )+
  scale_x_continuous(
    trans = ggallin::ssqrt_trans,
    labels = scales::comma,
    breaks = c(-100, 0, 25, 500),
    name = "Gens. after pathogen intro."
  )+
  scale_y_continuous(
    labels = scales::percent,
    breaks = c(0, 0.5, 1),
    name = "% Population infected"
  )+
  scale_fill_continuous_sequential(
    palette = "Reds 2"
  )+
  coord_cartesian(
    expand = T,
    xlim = c(sgen, genmax) - sgen,
    ylim = c(0, popsize) / popsize
  )+
  theme_test(
    base_size = 8
  )+
  theme(
    axis.text.y = element_text(
      angle = 90,
      hjust = 0.5,
      size = 6
    )
  )
```

## Wrap network and SIR figures for Fig. 3

```{r}
# wrap plots
fig_networks_disease =
  wrap_plots(
    wrap_plots(networkplots, guides = "collect"),
    fig_gen_infect, fig_sir,
    ncol = 2,
    design = "AAA\nAAA\nBCC"
  ) &
  plot_annotation(
    tag_levels = c("A")
  ) &
  theme(
    plot.tag = element_text(
      face = "bold"
    ),
    legend.position = "bottom"
  )

ggsave(
  plot = fig_networks_disease,
  filename = "figures/fig_networks_disease.png",
  height = 120,
  width = 120,
  units = "mm"
)
```

## Load networks from alternative scenarios

Read parameter files to subset for default parameter combination, based on the "ofi" column.

```{r}
params = list.files(
  "data/parameters", pattern = "csv", full.names = T
)
params = lapply(params, fread) |>
  rbindlist()
```

```{r}
params = params[(costInfect == 0.05 & regen_time == 50) | 
                  (costInfect == 0.25 & regen_time == 50 & dispersal == 10)]

data_files = glue("data/results/networks/data_networks_{params$ofi}.Rds")
```

```{r}
data_ntwk = lapply(data_files, function(file) {
  nt = readRDS(file)
  nt # we previously saved networks under the name nt
})
```

## Degree distribution in alternative scenarios

```{r}
degree_data = lapply(
  data_ntwk, function(le) {
    le_pre = le[["2500"]]
    le_post = le[["3500"]]
    # pre pathogen degree
    le_pre = mutate(
      le_pre, degree = tidygraph::centrality_degree(
        normalized = F
      ),
      type = "pre"
    ) |>
      as_tibble()
    
    # post pathogen degree
    le_post = mutate(
      le_post, degree = tidygraph::centrality_degree(
        normalized = F
      ),
      type = "post"
    ) |>
      as_tibble()
    
    rbindlist(
      list(
        le_pre,
        le_post
      )
    )
  }
) |>
  rbindlist()

# save degree data
fwrite(
  degree_data,
  "data/results/data_alt_degree_distribution.csv"
)
```

```{r}
degree_data = fread("data/results/data_alt_degree_distribution.csv")
degree_data$type = factor(degree_data$type, levels = c("pre", "post"))

# add simulation type
degree_data[, sim_type := fcase(
  infect_percent == "percent", "percent",
  dispersal == 10, "global"
)]
```

```{r}
# popsize per sim
popsize = 500
# how many replicates
replicates = 10

fig_degree_alt =
ggplot(degree_data)+
  geom_histogram(
    aes(
      degree,
      fill = type,
      y = ..count.. / (popsize * replicates)
    ),
    bins = 15,
    col = NA,
    show.legend = F
  )+
  geom_vline(
    xintercept = 0.1 * popsize,
    lty = 2,
    size = 0.2
  )+
  scale_fill_discrete_diverging(
    palette = "Blue-Red",
    rev = F,
    l1 = 50
  )+
  scale_x_continuous(
    trans = ggallin::pseudolog10_trans,
    breaks = c(0, 10, 100, 500),
    labels = function(x) {
      scales::percent(
        as.numeric(x) / popsize
      )
    },
    name = "% Pop. encountered"
  )+
  scale_y_continuous(
    # breaks = c(0, 0.1, 0.25),
    labels = scales::percent
  )+
  facet_grid(
    sim_type ~ type,
    labeller = labeller(
      type = c(
        post = "Post-pathogen", 
        pre = "Pre-pathogen"
      ),
      sim_type = c(
        global = "Global dispersal",
        percent = "Percentage infection cost"
      )
    )
  )+
  theme_test(
    base_family = "Arial",
    base_size = 8
  )+
  theme(
    legend.position = "top",
    legend.key.height = unit(0.5, "mm"),
    legend.key.width = unit(2, "mm"),
    axis.text.y = element_text(
      angle = 90,
      hjust = 0.5
    ),
    panel.background = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(
      face = "bold"
    )
  )+
  labs(
    y = "% Indiv."
  )+
  guides(
    fill = guide_legend()
  )
```

```{r}
ggsave(
  fig_degree_alt,
  filename = "supplement/figures/fig_degree_alt.png",
  height = 100,
  width = 120, units = "mm"
)
```
