---
editor_options: 
  chunk_output_type: console
---

## Figures for the vertical transmission scenario

```{r}
library(data.table)

library(ggplot2)
library(patchwork)
library(colorspace)
```

Load generation strategy data.

```{r}
files <- list.files(
  "data/results/morph_data/",
  pattern = "data_gen_vertical",
  full.names = TRUE
)

data_all <- lapply(files, fread)
data_all <- rbindlist(data_all)
# make a copy
df_strat <- copy(data_all)

# popsize
popsize <- 500
df_strat[, prop := N / popsize]

# get time since pathogen
sgen <- 500
genmax <- 1000
df_strat[, gen_abs := gen - sgen]

# save for plotting and upload
fwrite(
  df_strat,
  file = "data/results/data_vertical_evo.csv"
)
```

```{r}
# load saved data
df_strat <- fread("data/results/data_vertical_evo.csv")
```

## Plot evolutionary change across scenarios

```{r}
# fig_evo_social_strategy =
p_compare = 
  ggplot(df_strat[gen %% 2 == 0 & (social_strat != "non-handler tracking"), ]) +
  geom_path(
    aes(
      gen_abs, prop,
      colour = social_strat,
      group = interaction(repl, social_strat)
    ),
    size = 0.2
  ) +
  geom_vline(
    xintercept = c(0),
    lty = 2,
    size = 0.3,
    col = c("red")
  ) +
  facet_grid(
    cols = vars(p_v_transmit),
    rows = vars(cost),
    as.table = F,
    switch = c("y"),
    labeller = labeller(
      cost = function(x) sprintf("δE = %s", x),
      p_v_transmit = function(x) sprintf("p(V. Trans.) = %s", x)
    )
  ) +
  scale_colour_discrete_sequential(
    palette = "Viridis",
    l2 = 80,
    rev = FALSE,
    name = NULL,
    limits = c("agent avoiding", "agent tracking", "handler tracking"),
    labels = stringr::str_to_sentence,
    na.value = "lightgrey"
  ) +
  scale_x_continuous(
    # trans = ggallin::pseudolog10_trans,
    breaks = c(-25, 0, 25, 100, 200),
    labels = scales::comma_format(
      accuracy = 1
    ),
    name = "Gens. after pathogen intro.",
    sec.axis = sec_axis(
      trans = function(x) x + sgen,
      breaks = c(1, sgen, sgen + 100, sgen + 200),
      labels = scales::comma_format(
        accuracy = 1
      ),
      name = "Generations"
    )
  ) +
  scale_y_continuous(
    breaks = c(0, 0.5, 1),
    labels = scales::percent,
    name = "% Individuals"
  ) +
  coord_cartesian(
    xlim = c(-30, 250),
    ylim = c(0, 1),
    expand = FALSE
  ) +
  theme_test(
    base_family = "Arial",
    base_size = 8
  ) +
  theme(
    legend.position = "top",
    legend.key.height = unit(1, "mm"),
    legend.key.width = unit(2, "mm"),
    axis.text.x = element_text(hjust = 0.5),
    axis.text.y = element_text(
      angle = 90, hjust = 0.5
    )
  )

# save for supplement
ggsave(
  p_compare,
  filename = "supplement/figures/fig_evo_change_vertical.png",
  height = 120,
  width = 120,
  units = "mm"
)
```

## Plot all infections

```{r}
# read in infections data
df_vertical_infections = fread("data/results/data_infections_gen.csv")[
  scenario_tag == "vertical",
]
df_vertical_infections[, gen_abs := gen - sgen]

df_vertical_infections[, pathogen_eliminated := any(
  n_infected[(gen > sgen) & (gen < sgen + 200)] == 0
  ),
  by = c("cost", "p_v_transmit", "repl")]

# infection persistence
p_infections = ggplot(df_vertical_infections) +
  geom_path(
    aes(gen_abs, n_infected / popsize,
        group = repl, col = pathogen_eliminated),
    size = 0.2
  ) +
  geom_vline(
    xintercept = c(0),
    lty = 2,
    size = 0.3,
    col = c("red")
  ) +
  facet_grid(
    cols = vars(p_v_transmit),
    rows = vars(cost),
    as.table = F,
    switch = c("y"),
    labeller = labeller(
      cost = function(x) sprintf("δE = %s", x),
      p_v_transmit = function(x) sprintf("p(V. Trans.) = %s", x)
    )
  ) +
  scale_colour_discrete_sequential(
    palette = "Red-Yellow",
    l2 = 80,
    rev = FALSE, name = NULL,
    labels = c(
      "Pathogen persists", "Pathogen eliminated"
    ),
    na.value = "lightgrey",
  ) +
  scale_x_continuous(
    # trans = ggallin::pseudolog10_trans,
    breaks = c(-25, 0, 25, 100),
    labels = scales::comma_format(
      accuracy = 1
    ),
    name = "Gens. after pathogen intro.",
    sec.axis = sec_axis(
      trans = function(x) x + sgen,
      breaks = c(sgen, sgen + 100, sgen + 200),
      labels = scales::comma_format(
        accuracy = 1
      ),
      name = "Generations"
    )
  ) +
  scale_y_continuous(
    breaks = c(0, 0.5, 1),
    labels = scales::percent,
    name = "% Individuals"
  ) +
  coord_cartesian(
    xlim = c(-30, 250),
    ylim = c(0, 1)
  ) +
  theme_test(
    base_family = "Arial",
    base_size = 8
  ) +
  theme(
    legend.position = "top",
    legend.key.height = unit(1, "mm"),
    legend.key.width = unit(2, "mm"),
    axis.text.x = element_text(hjust = 0.5),
    axis.text.y = element_text(
      angle = 90, hjust = 0.5
    )
  )

# save for supplement
ggsave(
  p_infections,
  filename = "supplement/figures/fig_infection_vertical.png",
  height = 120,
  width = 120,
  units = "mm"
)
```

## Plot evolutionary change and infections

Pick two focal simulations in the cost = 0.25, p(v transmit) = 0.2 scenario.

```{r}
df_focus = df_strat[p_v_transmit == 0.2 & cost == 0.25, ]

# subset for parameters
df_vertical_infections = df_vertical_infections[p_v_transmit == 0.2 &
                                                  cost == 0.25, ]

# get sims where pathogen is eliminated
df_vertical_infections[scenario_tag == "vertical", pathogen_eliminated := any(
  n_infected[(gen > 500) & (gen < 700)] == 0
  ),
  by = c("cost", "p_v_transmit", "repl")]

df_vertical_infections = df_vertical_infections[
  repl == first(repl[pathogen_eliminated]) |
    repl == first(repl[!pathogen_eliminated])]

df_focus = df_focus[repl %in% df_vertical_infections$repl, ]
```

```{r}
p_evo = ggplot() +
  geom_col(
    data = df_focus,
    aes(gen_abs, prop, fill = social_strat),
    width = 2
  ) +
  geom_line(
    data = df_vertical_infections,
    aes(gen_abs, n_infected / popsize, col = "infected")
  ) +
  geom_vline(
    xintercept = c(0),
    lty = 2,
    size = 0.3,
    col = c("red")
  ) +
  scale_fill_discrete_sequential(
    palette = "Viridis",
    l2 = 80, alpha = 0.6,
    rev = FALSE,
    name = NULL,
    limits = c("agent avoiding", "agent tracking", "handler tracking"),
    labels = stringr::str_to_sentence,
    na.value = "lightgrey"
  ) +
  scale_colour_manual(
    values = c(infected = "grey20"),
    labels = "% Infected",
    name = NULL
  ) +
  facet_grid(
    rows = vars(repl),
    labeller = labeller(
      repl = c(
        "1" = "Pathogen persistance",
        "2" = "Pathogen eliminated"
      )
    ),
    switch = c("y")
  ) +
  scale_x_continuous(
    # trans = ggallin::pseudolog10_trans,
    breaks = c(-25, 0, 25, 100, 250, genmax),
    labels = scales::comma_format(
      accuracy = 1
    ),
    name = "Gens. after pathogen intro.",
    sec.axis = sec_axis(
      trans = function(x) x + sgen,
      breaks = c(1, sgen, sgen + 100, sgen + 250, genmax),
      labels = scales::comma_format(
        accuracy = 1
      ),
      name = "Generations"
    )
  ) +
  scale_y_continuous(
    breaks = c(0, 0.5, 1),
    labels = scales::percent,
    name = "% Individuals"
  ) +
  coord_cartesian(
    xlim = c(-30, 150),
    ylim = c(0, 1),
    expand = FALSE
  ) +
  theme_test(
    base_family = "Arial",
    base_size = 8
  ) +
  theme(
    legend.position = "top",
    legend.key.height = unit(1, "mm"),
    legend.key.width = unit(2, "mm"),
    axis.text.x = element_text(hjust = 0.5),
    axis.text.y = element_text(
      angle = 90, hjust = 0.5
    )
  )
```

## Social information importance

```{r}
files <- list.files(
  "data/results/si_imp_data",
  pattern = "vertical",
  full.names = TRUE
)

data_all <- lapply(files, fread)
data_all <- rbindlist(data_all)

# make a copy
df_si = copy(data_all)

df_si <- df_si[(cost == 0.25) & (p_v_transmit == 0.2)]

# popsize
popsize <- 500

df_si[, prop := N / popsize]

# get time since pathogen
sgen <- 500
genmax <- 1000

# save for plotting and upload
fwrite(
  df_si,
  file = "data/results/data_vertical_default_si_imp.csv"
)
```

```{r}
# load saved data
df_si <- fread("data/results/data_vertical_default_si_imp.csv")
```

```{r}
# get focal replicates
df_si = df_si[repl %in% unique(df_focus$repl)]
df_si[, si_imp := plyr::round_any(si_imp, accuracy = 0.01)]
df_si = df_si[, list(prop = sum(prop),
                     N = sum(N)), by = c("gen", "repl", "si_imp")]
df_si[, gen_abs := gen - sgen]
```


```{r}
# get weighted mean and variance in social info importance
# replicate 2 is an anomaly with very high SI use, although strategies are
# similar across replicates
df_si_mean <- df_si[, list(
  mean = weighted.mean(si_imp, w = N),
  sd = Hmisc::wtd.var(si_imp, weights = N)
), by = c("gen", "repl")]

df_si_mean[, c("y", "ymin", "ymax") := list(
  mean,
  mean - sd,
  mean + sd
)]
df_si_mean[, gen_abs := gen - sgen]
```

### Figure social information importance

```{r}
hmap_cols <- colorspace::sequential_hcl(
  5,
  palette = "sunset",
  rev = TRUE
)

p_si_importance =
ggplot() +
  geom_tile(
    data = df_si, # 7
    aes(
      gen_abs, si_imp,
      fill = N / popsize
    ),
    width = 2
  ) +
  facet_grid(
    rows = vars(repl),
    labeller = labeller(
      repl = c(
        "1" = "Pathogen persistance",
        "2" = "Pathogen eliminated"
      )
    ),
    switch = c("y")
  ) +
  geom_pointrange(
    data = df_si_mean[gen %% 20 == 0,],
    aes(
      gen_abs, y,
      ymin = ymin,
      ymax = ymax
    ),
    size = 0.1, shape = 1,
    colour = "grey40"
  ) +
  geom_hline(
    yintercept = 0.5,
    col = "grey",
    lty = 2,
    size = 0.2
  ) +
  geom_vline(
    xintercept = c(0),
    lty = 2,
    size = 0.3,
    col = c("red")
  ) +
  scale_fill_gradientn(
    colours = c(
      "grey99",
      hmap_cols
    ),
    na.value = hmap_cols[5],
    limits = c(0.0, 0.3),
    breaks = c(0.01, 0.3, 0.5),
    labels = function(x) {
      a <- scales::percent(x)
      a[x == 0.3] <- ">30%"
      a
    },
    name = "% Indiv."
  ) +
  scale_x_continuous(
    # trans = ggallin::pseudolog10_trans,
    breaks = c(0, 100, 250),
    labels = scales::comma_format(
      accuracy = 1
    ),
    name = "Gens. after pathogen intro.",
    sec.axis = sec_axis(
      trans = function(x) x + sgen,
      labels = scales::comma_format(
        accuracy = 1
      ),
      name = "Generations"
    )
  ) +
  scale_y_continuous(
    labels = scales::percent,
    name = "% Contribution of social information"
  ) +
  coord_cartesian(
    xlim = c(-50, NA),
    ylim = c(0, 0.75)
  ) +
  theme_test(
    base_family = "Arial",
    base_size = 8
  ) +
  theme(
    legend.position = "top",
    legend.key.height = unit(1, "mm"),
    legend.key.width = unit(2.5, "mm"),
    legend.title = element_text(
      vjust = 1.5
    ),
    legend.text = element_text(
      size = 6
    ),
    # axis.text.x = element_text(hjust = 0.5, size = 6),
    axis.text.y = element_text(
      angle = 90, hjust = 0.5
    )
  )
```

## Ecological outcomes

Load generation data.

```{r}
# list files, read, and filter for default model options
files <- list.files(
  "data/results/gen_data",
  pattern = "vertical",
  full.names = TRUE
)

data_all <- lapply(files, fread)
data_all <- rbindlist(data_all)

df_eco <- data_all[cost == 0.25 & p_v_transmit == 0.2,]

# popsize
popsize <- 500

# get time since pathogen
sgen <- 500
genmax <- 1000
df_eco[, gen_abs := gen - sgen]

# save for plotting and upload
fwrite(
  df_eco,
  file = "data/results/data_vertical_default_eco.csv"
)
```

```{r}
# load saved data
df_eco <- fread("data/results/data_vertical_default_eco.csv")
```

```{r}
# subset for focal replicates
df_eco <- df_eco[repl %in% df_focus$repl, c(
  "repl", "gen_abs", "intake.mean",
  "moved.mean", "assoc.mean"
)]
df_eco <- melt(
  df_eco,
  id.vars = c("gen_abs", "repl")
)

# convert to percentage
df_eco[, value_scaled := value / value[gen_abs == 0], by = c(
  "repl", "variable"
)]

## colours for background
cols <- colorspace::diverging_hcl(3, palette = "Tofino", l = 50, c = 80)
```

```{r}
p_associations =
ggplot(df_eco[variable == "assoc.mean"]) +
  stat_summary(
    aes(
      gen_abs, value,
      col = "Associations"
    ),
    show.legend = FALSE,
    geom = "line"
  ) +
  scale_y_continuous(
    trans = "log10",
    labels = scales::comma
  ) +
  facet_grid(
    rows = vars(repl),
    labeller = labeller(
      repl = c(
        "1" = "Pathogen persistance",
        "2" = "Pathogen eliminated"
      )
    ),
    switch = c("y")
  ) +
  geom_vline(
    xintercept = c(0),
    lty = 2,
    size = 0.3,
    col = c("red")
  ) +
  scale_colour_manual(
    values = c(
      assoc.mean = cols[2]
    )
  ) +
  scale_x_continuous(
    # trans = ggallin::pseudolog10_trans,
    breaks = c(-30, 0, 100),
    labels = scales::comma_format(
      accuracy = 1
    ),
    name = "Gens. after pathogen intro.",
    sec.axis = sec_axis(
      trans = function(x) x + sgen,
      breaks = c(1, sgen, genmax),
      labels = scales::comma_format(
        accuracy = 1
      ),
      name = "Generations"
    )
  ) +
  coord_cartesian(
    xlim = c(-30, 150)
  ) +
  theme_test(
    base_family = "Arial",
    base_size = 8
  ) +
  theme(
    legend.position = "top",
    legend.key.height = unit(1, "mm"),
    legend.key.width = unit(2, "mm"),
    axis.text.x = element_text(hjust = 0.5),
    axis.text.y = element_text(
      angle = 90, hjust = 0.5
    )
  ) +
  labs(
    x = NULL,
    y = "Associations",
    colour = NULL,
    shape = NULL
  )
```

```{r}
fig_eco_evo_vertical <-
  wrap_plots(
    p_evo,
    p_si_importance,
    p_associations,
    ncol = 3,
    guides = "collect"
  ) &
    plot_annotation(
      tag_levels = c("A", 1)
    ) &
    theme(
      legend.position = "bottom",
      legend.justification = "left",
      # legend.key.height = unit(1, units = "mm"),
      plot.tag = element_text(
        face = "bold",
        size = 10
      ),
      axis.text.x = element_text(
        size = 7
      )
    )

ggsave(
  fig_eco_evo_vertical,
  filename = "figures/fig_eco_evo_vertical.png",
  height = 120, width = 180, units = "mm"
)
```
